Drivers of tree regeneration remain challenging to capture for models of forest dynamics

### 2022-11-17 manuscript/manuscriptRW.Rmd
added protocol information and model section in manuscript material and methods

---
title: "Capturing regeneration in models of forest dynamics"
author:
- Olalla Díaz-Yáñez, Forest Ecology ETH Zurich
- Yannek Käber, Forest Ecology ETH Zurich
- Mats Mahnken, PIK Potsdam
- Samuel Fischer, UFZ
- Rico Fischer, UFZ
- Friedrich Bohn, UFZ
- Xavier Morin, cefe cnrs
- Manfred Lexer, BOKU
- Christian Hochauer, BOKU
- Katarina Merganicova, forim
- Tobias Mette, LWF Bayern
- Werner Rammer, TUM
- Kristin H. Braziunas, TUM
- Paola Mairota, UNIBA
- Josef Bruna, ibot cas
- Heike Lischke, WSL
- Daniel Scherrer, nWSL
- Simon Scheiter, senckenberg
- Thomas Hickler, senckenberg
- Tim Anders, senckenberg
- Jessica Hetzer, senckenberg
- Harald Bugmann, Forest Ecology, ETH Zurich
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

In comparison to recruitment, growth and mortality responses to climate
change are much more studied

@@

A total of 15 models and model variants of forest dynamics at the stand,
landscape and global level were used to simulate stand dynamics and
sample regeneration in 200 sites selected across Europe.

About the models The models ( Table @ref(tab:tableModels)) were created
with different philosophies to represent regeneration. Three models,
LPJ-GUESS, the two variants of ForClim and Treemig simulated more
species than the one asked because XXXX , these species were assigned to
"other" category in the case of TreeMig and in the case of LPJ-GUESS
they were removed from the outputs as they are shrubs species (Boreal
evergreen shrub, Juniperus oxycedrus and quercus coccifera) LPJ-GUESS
also simulate Corylus avellana but only had 3 trees across all samples
and they were removed because due to the low degrees of freedom it was
not possible to create a model for the category "others" for this
model). Forclim simulated a total of 17 species more than the species
asked in the protocol . Observed data also have a category others.

About the sites The sites represent the available range of environmental
gradients in temperature and precipitation (Figure using the variables
degree day and water balance. The sites were selected across a network
of plots that collect information on forest reserves
[@kaber_sheltered_nodate]

Each modeler group was provided with detailed information on how to
perform the simulations including the input variables such as soil or
climate, simulation setup and expected output variables. The overarching
goal of this simulations was the assessment of the regeneration niche as
it arises from empirical data analysis vs. its representation in a wide
range of models of forest dynamics.

@@

@@

modelDF$Type[modelDF$Type ==  "Stand"] <- "S"
modelDF$Type[modelDF$Type ==  "Landscape"] <- "L"
modelDF$Type[modelDF$Type ==  "Global"] <- "G"

@@

modelDF$Reference <- "reference"

kable(modelDF,  format = "markdown", caption = "List of models included.")
```

@@

>>>>

---
title: "Drivers of tree regeneration remain challenging to capture for models of forest dynamics"
author:
- Olalla Díaz-Yáñez, Forest Ecology ETH Zurich
- Yannek Käber, Forest Ecology ETH Zurich
- Mats Mahnken, PIK Potsdam
- Samuel Fischer, UFZ
- Rico Fischer, UFZ
- Friedrich Bohn, UFZ
- Xavier Morin, cefe cnrs
- Manfred Lexer, BOKU
- Christian Hochauer, BOKU
- Katarina Merganicova, forim
- Tobias Mette, LWF Bayern
- Werner Rammer, TUM
- Kristin H. Braziunas, TUM
- Paola Mairota, UNIBA
- Josef Bruna, ibot cas
- Heike Lischke, WSL
- Daniel Scherrer, nWSL
- Simon Scheiter, senckenberg
- Thomas Hickler, senckenberg
- Tim Anders, senckenberg
- Jessica Hetzer, senckenberg
- Harald Bugmann, Forest Ecology, ETH Zurich
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

In comparison to recruitment, growth and mortality [@bugmann2019]
responses to climate change are much more studied

@@

## The models

Fifteen models of forest dynamics (including two model variants) were
used to assess tree regeneration. The models varied on their scale
including models at the stand, landscape and global scale (see Table
\@ref(tab:tableModels)). Their model construction philosophies and
origin also differs with two models where their formulations were
derived using statistical analyses based on inventory data (SYBILA and
xComp) and others with largely process based approach. Some of the
models largely rely on the original approach underlying forest gap
models (ForClim 1, ForClim 11, LandClim, Landis II ); and some include
plant ecophysiology formulations (FORMIND, PICUS, 4C, iLand). This
different approaches in the models construction also has influence how
regeneration is defined.

The models can be categorized as regeneration models and recruitment (or
ingrowth) models [@vanclay1994, @konig_tree_2022]. Regeneration models
include in their regeneration modules processes such as flowering and
pollination, seed production, seed dispersal, germination or seedling
establishment that lead to the number of seedlings or saplings.
Recruitment models introduce a number of seedlings or saplings with
certain characteristic (biomass or dbh) without considering earlier
development processes. We can also distinguish models that have feedback
in the regeneration as those in which the level and species composition
in regeneration is influenced by the existing adult tree community, this
can be done via the production of seed, seedlings or saplings. The
starting point for regeneration it was also different per model varying
from seedbank, seed, sapling and ingrowth. The overall complexity in the
regeneration modules also varied, following @bugmann2022 we can see that
most of the models have rather low/high? complexity while few of them
have high/low?. Further details on the models specific formulation
related to their regeneration modules and an overall summary across
models is presented in the supplementary materials.

e used repeated tree census data collected in forest reserves and
unmanaged (or protected) forests across Europe, 127 which were compiled
in the framework of the EuFoRIa network.

The sampled forests have not been managed for 128 at least one decade,
but most for much longer, and some are primary forests.

The final data set contained 701 units, each covering on average 1 ha of
forest area with 142 observations of tree recruitment at a DBH of 7 cm
(see

aggregated

## The observed data

Regeneration observations for a wide range of environmental gradients
are hard to obtain and this is one of the reasons why most of the models
of forest dynamics have never been confronted with a data-set covering a
wide range of regeneration sites across climate gradients to evaluate
how well are they capturing regeneration.

The observed data represent the available range of environmental
gradients in temperature and precipitation from 200 sites of strict
forest reserves in Europe compiled in the framework of the EuFoRIa
network. The sample forest have not been managed in at least 10 years,
and each of the sites contains at least two consecutive measurements.
Regeneration was measured at 7 and 10 cm threshold, however 35 sites
only contain information for 10 cm diameter. The 200 sites represent the
available climatic gradients of temperature and precipitation across
Europe and they varied in species composition (Figure \@ref(fig:map)).
For more information on this dataset and their regeneration please refer
to @kaber_sheltered_nodate.

## Simulation protocol

Each modeler group was provided with detailed protocol (available in the
supplementary material) on how to perform the simulations regarding
input variables such as climate or soil, simulation setup and expected
output variables. No information was provided on the regeneration on
these sites before the models performed were done. The overarching goal
of this simulations was the assessment of the regeneration niche as it
arises from empirical data analysis against its representation in a wide
range of models of forest dynamics.

The input variables were collected and aggregated to be adapted to the
different model demands. The time series of climatic variables from 1981
to 2018 were provided in hourly (Era5-land data @muñozsabater ), daily
and monthly resolution (CHELSA data, @karger ). Some variables required
for some models such as relative humidity or vapor pressure deficit had
to be calculated from these variables. The final use of the climate data
was different in each model (Table \@ref(tab:tableModels))The soil
quality data was provided as a continuous value between 1 and 5
(Soilgrids dataset, @hengl2017, @poggio2021 ). Elevation, slope and
aspect was also provided for each of the 200 sites, however no other
spatial information such as coordinates were provided (@astergd ).

The simulations had to be run in the absence of natural disturbances,
however one the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE
[ONE SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT
TYPE. The simulations had to be set up in a way that they will obtain
samples on species specific regeneration rates per decade and per ha.
The modeling teams decided on the simulated area and how they derive
these samples (further details on how each modeling team prepared the
simulations are available in the supplementary material).

The simulations were run from bare ground in the absence of management
to a simulated equilibrium (\"potential natural vegetation\") with the
current climate. The goal was to derive the regeneration niche.The
regeneration niche is defined as the passing of a diameter threshold of
7 and 10 cm (ingrowth; cf. below). Both the width of the regeneration
niche (i.e., in environmental space) as well as the intensity of the
recruitment process (i.e., the number of ingrowth trees per area and per
unit of time) are of interest. The exact length of the simulation should
was decided by the modeling teams (Table \@ref(tab:tableModels)).

The simulations were run for mixed species forests (not multiple
single-species simulations) using mixtures of the following
species/genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus
betulus, Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus
excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris.* Three
models included additional species in the simulations (ForClim 1,
ForClim 11 and TreeMig) and three could not simulate all the species
(4C, xComp, TreeMig, LPJ-GUESS). HERE **FORCLIM** PROVIDES [ONE
SENTENCE]{.underline} WHY IT SIMULATED MORE SPECIES. HERE **TREEMIG**
PROVIDES [ONE SENTENCE]{.underline} WHY IT SIMULATED MORE SPECIES AND
MISSED ONE. HERE **4C** PROVIDES [ONE SENTENCE]{.underline} WHY IT
SIMULATED LESS SPECIES. HERE **XCOMP** PROVIDES [ONE
SENTENCE]{.underline} WHY IT SIMULATED LESS SPECIES. The model aDGVM2
did not have species level option and provided the default approach with

The simualtion have to be provided for two diameter thresholds: 7 and 10
cm.

@@

```{r map, echo=FALSE, fig.cap="Map with the site observations and species presnet in regeneration.", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

## Analysis

@@

# modelDF$Type[modelDF$Type ==  "Stand"] <- "S"
# modelDF$Type[modelDF$Type ==  "Landscape"] <- "L"
# modelDF$Type[modelDF$Type ==  "Global"] <- "G"

@@

kable(modelDF, 
      #format = "markdown",
      format = "latex", booktabs = TRUE,
      caption = "List of models included.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |> 
      kableExtra::column_spec(column = 2, width = "5em") |> 
      kableExtra::column_spec(column = 3, width = "5em") |> 
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "8em") |>
      kableExtra::column_spec(column = 8, width = "5em") |>
      kableExtra::column_spec(column = 9, width = "5em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "5em") 
```

@@

-    Generally, the approach was appreciated a lot, i.e. the idea of
    confronting the simulated recruitment along environmental gradients
    with data from such gradients. With EuFoRIa, we seem to be on a very
    good track in this regard.\
    \
    • Critical comments were given to the choice of an ingrowth limit of
    10 cm DBH. Many participants said, in one way or the other, that the
    good match is not surprising because if the model gets overall
    species composition right, it must also get ingrowth of these rather
    large trees right. Thus, a better test would be for the ingrowth
    limit to be lower. My explanation that we wanted to maximize the
    dataset available and thus chose the highest callipering limit was
    accepted, but I think it is definitely so that we need to provide
    data with a lower limit. How much data do we lose if we go down to 4
    cm? I am afraid that going down to 1 cm is not realistic because
    we\'d lose most of the data. So the bottomline is that 10 cm are
    fine, but a very soft test, and hence we need to provide material
    for a harder test. Can you tell me (1) how much data we\'d lose with
    4 cm, and (2) what the patterns along the two gradients (CWB and DD)
    are looking like when using a 4 cm threshold?\
    \
    • Many participants also felt that we should not miss out on a third
    very important axis, i.e. light availability. Obviously this is more
    tricky because we do not have such measurements, but I think we can
    work with proxies such as basal area (although it\'s difficult for
    the small sampling plots, as light availability may be poorly
    related to BA on a 500 m2 plot...). Have you looked into this
    already? It would be really nice if we could provide a pattern that
    shows how recruitment of the different species is becoming sparser
    and drops out completely as a function of a proxy of light
    availability.\
    \
    Beyond that, some people expressed their desire to calibrate a
    recruitment model based on the EuFoRIa data, and I tried to explain
    why I think it is more insightful to use these data for model
    benchmarking: we can learn more... :-)\
